---
title: "C2W4 ProgAssignment 1-2"
author: "Jeff"
date: "May 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Plot the 30-day mortality rates for heart attack
Read the outcome data into R via the read.csv function and look at the rst few rows.
> outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
> head(outcome)

```{r}
outcome <- read.csv("~/R Works/Course 2/rprog_data_ProgAssignment3-data/outcome-of-care-measures.csv", colClasses = "character") #is it necessory to read all col's as character?
head(outcome,3)
```
```{r}
dim(outcome)
```
```{r}
names(outcome)
```

```{r}
outcome[,11]<-as.numeric(outcome[,11])  #coerce to numeric for hist() to work
## You may get a warning about NAs being introduced; that is okay
hist(outcome[,11])
```

To understand the tall bar to the left
```{r}
summary(outcome[,11])
a<-outcome[outcome[,11]>90,11]
summary(a)
sum(outcome[,11]==96)
```
```{r}
unique(outcome[,7])
```
## 2 Finding the best hospital in a state

Write a function called best that take two arguments: the 2-character abbreviated name of a state and an
outcome name. The function reads the outcome-of-care-measures.csv le and returns a character vector
with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the speci ed outcome
in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can
be one of \heart attack", \heart failure", or \pneumonia". Hospitals that do not have data on a particular
outcome should be excluded from the set of hospitals when deciding the rankings.
Handling ties. If there is a tie for the best hospital for a given outcome, then the hospital names should
be sorted in alphabetical order and the rst hospital in that set should be chosen (i.e. if hospitals \b", \c",
and \f" are tied for best, then hospital \b" should be returned).
The function should use the following template.
best <- function(state, outcome) {
## Read outcome data
## Check that state and outcome are valid
## Return hospital name in that state with lowest 30-day death
## rate
}
The function should check the validity of its arguments. If an invalid state value is passed to best, the
function should throw an error via the stop function with the exact message \invalid state". If an invalid
outcome value is passed to best, the function should throw an error via the stop function with the exact
message \invalid outcome".

```{r}
best <- function(state, outcome) {
        ## Read outcome data
        outcome_data <- read.csv("~/R Works/Course 2/rprog_data_ProgAssignment3-data/outcome-of-care-measures.csv", colClasses = "character") 
        
        ## Check that state and outcome are valid
        st<-unique(outcome_data[,7])
        if (!is.element(state,st)) return("invalid state")
        if (!is.element(outcome,c("heart attack","heart failure","pneumonia"))) return("invalid outcome")
        
        ## Return hospital name in that state with lowest 30-day death rate
        col<-c("heart attack"=11,"heart failure"=17,"pneumonia"=23)
        st_data<-outcome_data[outcome_data[,7]==state & outcome_data[,col[outcome]]!="Not Available",c(2,col[outcome])]
        st_data[,2]<-as.numeric(st_data[,2])         # It is important to convert to numeric here.          
        best_hos<-st_data[which(st_data[,2]==min(st_data[,2],na.rm = TRUE)),1]
        sort(best_hos)
        best_hos[1]
        
}

best("TX", "heart attack")

```

```{r}
best("TX", "heart failure")
```

```{r}
best("MD", "heart attack")
```

```{r}
best("MD", "pneumonia")
```

```{r}
best("BB", "heart attack")
```

```{r}
best("NY", "hert attack")
```

```{r}
best("SC", "heart attack")
```
```{r}
best("NY", "pneumonia")
```


```{r}
best("AK", "pneumonia")
```
