---
title: "Course 2 Week 1 Quiz"
author: "Jeff"
date: "May 23, 2019"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Quiz 1
Q12. Read the first 2 rows
```{r}
##Read file to workspace
hw1_data<-read.csv(file.path("~","R Works","Course 2","hw1_data.csv"))
```   

```{r}
head(hw1_data,2) 
```    

13.number of observations in the data frame
```{r}        
nrow(hw1_data)   
```

        
14.last 2 rows
```{r}
tail(hw1_data,2) 
```        
        
15. Ozone value in 47th row
```{r}
hw1_data[47, "Ozone"]  
```      
        
16. count of missing values in Ozone
```{r}        
sum(is.na(hw1_data$Ozone)) 
```      
        
17. mean of the Ozone column in this dataset? Exclude missing values
```{r}
mean(na.omit(hw1_data$Ozone)) 
```
        
        
18.Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset?
```{r}
mean(hw1_data[which(hw1_data$Ozone>31 & hw1_data$Temp>90),"Solar.R"]) 
```     
        
19. What is the mean of 'Temp' when 'Month' is equal to 6?
```{r}
mean(hw1_data[which(hw1_data$Month == 6), "Temp"],na.rm = TRUE)
``` 
        
20. What was the maximum ozone value in the month of May?
```{r}
max(hw1_data[hw1_data$Month == 5 , "Ozone"], na.rm=TRUE) 
```
                